---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


### genpwr

The genpwr package for R (>3.5.1) performs power and sample size calculations for genetic association studies and allows for mis-specification of the genetic model.  Calculations can be performed for binary (case/control) and continuous outcomes.  Power and sample size calculations are possible for genetic effects as well as gene by environment interactions.

### Example

To calculate power to detect an odds ratio of 2 for a 1:1 case control study with 2,000 subjects, assuming an alpha of 0.05, at minor allele frequencies of 0.1, 0.2, and 0.3:

```{r}
library(genpwr)

genpwr.calc(calc = "power", model = "logistic", N = 2000, OR = 2,
            Alpha = 0.05, MAF = c(0.1,0.2,0.3), Case.Rate = 0.5)
```

"The return object contains information about power for additive, dominant, recessive, and 2df / genotypic tests of association, assuming various true underlying genetic effects (additive, dominant, recessive). "


### Installation instructions

The results in the above manuscript were produced using R version 3.5.1. To reproduce the results,
perform the following steps:

* Install R version 3.5.1 or higher by following the instructions at http://www.r-project.org
* From the R environment, install and load the "devtools" package
```R
install.packages("devtools")
library(devtools)
```
* Install the "genpwr" package directly from Github.com
```R
install_github("camillemmoore/Power_Genetics", subdir="genpwr")
```
* Load the library
```R
library(genpwr)
```

### Demo

Install the genpwr package as described above.

Run the genwr demo program

```R
demo(genpwr_demo)
```


